---
title: "NEWS"
author: Erik.Leppo@tetratech.com and Tim.Martin@state.mn.us
output: github_document
  #md_document:
   # variant: markdown_github
---

<!-- NEWS.md is generated from NEWS.Rmd. Please edit that file -->

```{r Options, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "NEWS-"
)
```

```{r LastUpdate, echo = FALSE}
cat(paste0("Last Update: ",Sys.time()))
```
LakeMonitoR 0.1.0.9012
===========
Released - 2021-01-19

* refactor: Remove docs folder
    + Using GitHub Action to build pkgdown site

LakeMonitoR 0.1.0.9011
===========
Released - 2021-01-19

* style: Remove blank lines at end of script
    + daily_depth_means
    + data
    + exort_rLakeAnalyzer
    + LakeMonitoR
    + runShiny
    + stratification
* style: Trim lines to 80 characters
    + runShiny
    + daily_depth_means
    + data
    + export_rLakeAnalyzer 
    + stratification
* refactor: Add items to .Rbuildignore and .gitignore
    + _other folder
    + NEWS.rmd
    + README.RMD
    + Switch setting to use gh-pages branch
* tests: Add two tests for stratification (dates and events)
* tests: Add test for export_rLakeAnalyzer
* docs: Add all extra export_rLakeAnalyzer examples to dontrun
    + Comment out failing schmidt.plot()
* refactor: Add missing parameters to function call
    + dir_export
    + fn_export
* docs: Update DESCRIPTION license
    
LakeMonitoR 0.1.0.9010
===========
Released - 2021-01-19

* tests: Add test for daily_depth_means

LakeMonitoR 0.1.0.9009
===========
Released - 2020-12-27

* docs: ReadMe, Add license badge
* docs: ReadMe, Add CodeFactor badge
* docs: ReadMe, Add lifecyle badge
* docs: ReadMe, Add maintenance badge
* docs: ReadMe, Add issues badge
* docs: ReadMe, Add release badge
* docs: ReadMe, Add download badge
* docs: ReadMe, Add codecov badge
* ci: Add GitHub Action, CI
* docs: ReadMe, Add CMD Check badge
* ci: Add GitHub Action, pkgdown
* docs: Remove docs folder from main branch
* docs: GitHub redirect web page to gh-pages branch
* ci: Add GitHub Action, test coverage
* docs: ReadMe, rework md as rmd

LakeMonitoR 0.1.0.9008
===========
Released - 2020-11-13
* server.R
    + Add inputs to sink.
    + Min days input wasn't getting used.
    + Min days input to numeric.
* stratification.R
    + Convert min_days to numeric as a QC check.

LakeMonitoR 0.1.0.9007
===========
Released - 2020-11-13
* stratification.R
    + Add number of days to column name in output, Issue #3.
* check
    + Add and declare packages for functions in DESCRIPTION and code.

LakeMonitoR 0.1.0.9006
===========
Released - 2020-11-10

* Add package help file.
    
LakeMonitoR 0.1.0.9005
===========
Released - 2020-10-27

* Shiny, ui.R
    + Update version number in header.

LakeMonitoR 0.1.0.9004
===========
Released - 2020-10-27

* Shiny, server.R
    + Add error checking for column names and upload files, Issue #1

LakeMonitoR 0.1.0.9003
===========
Released - 2020-10-26

* Update Shiny app.
* Add example files for Shiny app.

LakeMonitoR 0.1.0.9002
===========
Released - 2020-10-22

* ReadMe.md
    + Fix typo in  install_github example code.
* Add function export_rLakeAnalyzer
    + Convert data frame to rLakeAnalyzer input format.
* DESCRIPTION
    + Imports: reshape2
    + Suggests: rLakeAnalyzer and ggplot2
* Data.R
    + Add lakehypso

LakeMonitoR 0.1.0.9001
===========
Released - 2020-10-22

* Add Shiny app.
    + File upload only.
* runShiny.R
    + Add function to run shiny app.

LakeMonitoR 0.1.0
===========
Released - 2020-10-20

* Interim version.
    + Both functions fully operational.
* Update pkgdown.

LakeMonitoR 0.0.1.9006
===========
Released - 2020-10-20

* Remove docs (pkgdown) from gitignore

LakeMonitoR 0.0.1.9005
===========
Released - 2020-10-20

* Add pkgdown documentation.

LakeMonitoR 0.0.1.9004
===========
Released - 2020-10-20

* Update data-raw and data with full records.
    + Use full raw data.
    + Zip raw data file.
* Update data.R

LakeMonitoR 0.0.1.9003
===========
Released - 2020-10-19

* Add stratification function.

LakeMonitoR 0.0.1.9002
===========
Released - 2020-10-19

* Update daily_depth_means function.
* Add data file, laketemp_ddm (output of daily_depth_means function).

LakeMonitoR 0.0.1.9001
===========
Released - 2020-10-16

* Add laketemp data
* Add functions.
    + daily_depth_means
    + lake_strat

LakeMonitoR 0.0.1.9001
===========
Released - 2020-10-16

* Initial commit to GitHub
