---
title: "NEWS"
author: Erik.Leppo@tetratech.com and Tim.Martin@state.mn.us
output: github_document
  #md_document:
   # variant: markdown_github
---

<!-- NEWS.md is generated from NEWS.Rmd. Please edit that file -->

```{r Options, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "NEWS-"
)
```

```{r LastUpdate, echo = FALSE}
cat(paste0("Last Update: ",Sys.time()))
```
LakeMonitoR 0.1.0.9003
===========
Released - 2020-10-26
* Update Shiny app.
* Add example files for Shiny app.

LakeMonitoR 0.1.0.9002
===========
Released - 2020-10-22

* ReadMe.md
    + Fix typo in  install_github example code.
* Add function export_rLakeAnalyzer
    + Convert data frame to rLakeAnalyzer input format.
* DESCRIPTION
    + Imports: reshape2
    + Suggests: rLakeAnalyzer and ggplot2
* Data.R
    + Add lakehypso

LakeMonitoR 0.1.0.9001
===========
Released - 2020-10-22

* Add Shiny app.
    + File upload only.
* runShiny.R
    + Add function to run shiny app.

LakeMonitoR 0.1.0
===========
Released - 2020-10-20

* Interim version.
    + Both functions fully operational.
* Update pkgdown.

LakeMonitoR 0.0.1.9006
===========
Released - 2020-10-20

* Remove docs (pkgdown) from gitignore

LakeMonitoR 0.0.1.9005
===========
Released - 2020-10-20

* Add pkgdown documentation.

LakeMonitoR 0.0.1.9004
===========
Released - 2020-10-20

* Update data-raw and data with full records.
    + Use full raw data.
    + Zip raw data file.
* Update data.R

LakeMonitoR 0.0.1.9003
===========
Released - 2020-10-19

* Add stratification function.

LakeMonitoR 0.0.1.9002
===========
Released - 2020-10-19

* Update daily_depth_means function.
* Add data file, laketemp_ddm (output of daily_depth_means function).

LakeMonitoR 0.0.1.9001
===========
Released - 2020-10-16

* Add laketemp data
* Add functions.
    + daily_depth_means
    + lake_strat

LakeMonitoR 0.0.1.9001
===========
Released - 2020-10-16

* Initial commit to GitHub
