---
title: "LakeMonitoR Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rLakeAnalyzer Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE
                      , comment = "#>"
                      , fig.dim = c(8, 6)/1.5
                      , fig.show='asis'
                      #, out.width = "50%"
                      )
library(LakeMonitoR)
```

# Purpose
Show users how to use functions in `LakeMonitoR` and those from `rLakeAnalyzer`.

For rLakeAnalyzer examples, in most cases, are using examples provided by 
`rLakeAnalyzer`.

# LakeMonitoR

## agg_depth_files

```{r agg_depth_files, echo=TRUE, eval=FALSE}
# Data Files
myFile_import <- c("Ellis--1.0m_Water_20180524_20180918.csv"
                  , "Ellis--3.0m_Water_20180524_20180918.csv")
myFile_export <- "Ellis--Combined_Water_20180524_20180918.csv"
myDir_import <- file.path(system.file("extdata", package = "LakeMonitoR"))
myDir_export <- tempdir()
agg_depth_files(filename_import = myFile_import
                , filename_export = myFile_export
                , dir_import = myDir_import
                , dir_export = myDir_export)
```

## daily_depth_means

```{r daily_depth_means}
# Packages
library(xts)

# Lake Data
data <- laketemp

# Filter by any QC fields
data <- data[data$FlagV == "P", ]

# name columns
col_siteid   <- "SiteID"
col_datetime <- "Date_Time"
col_depth    <- "Depth_m"
col_measure  <- "Water_Temp_C"

# run function
data_ddm <- daily_depth_means(data
                             , col_siteid
                             , col_datetime
                             , col_depth
                             , col_measure)

# summary
summary(data_ddm)
```

## lake_summary_stats

```{r lake_summary_stats}
# data
data <- laketemp_ddm

# Columns
col_date    <- "Date"
col_depth   <- "Depth"
col_measure <- "Measurement"
below_threshold <- 2

# Calculate Stratification
df_lss <- lake_summary_stats(data
                            , col_date
                            , col_depth
                            , col_measure
                            , below_threshold)

# Results
head(df_lss)
```

## plot_depth

```{r plot_depth}
library(LakeMonitoR)

# Data (Test Lake)
data         <- laketest

# Column Names
col_datetime <- "Date.Time"
col_depth    <- "Depth"
col_measure  <- "temp_F"

# Plot Labels
lab_datetime <- "Date Time"
lab_depth    <- "Depth (m)"
lab_measure  <- "Temperature (F)"
lab_title    <- "Test Lake"

# Create Plot
p_profile <- plot_depth(data = data
                        , col_datetime = col_datetime
                        , col_depth = col_depth
                        , col_measure = col_measure
                        , lab_datetime = lab_datetime
                        , lab_depth = lab_depth
                        , lab_measure = lab_measure
                        , lab_title = lab_title)

# Print Plot
print(p_profile)
```

The function is generic and can plot other variables if present.

```{r plot_depth_do}
# Data (Test Lake)
data         <- laketest

# Column Names
col_datetime <- "Date.Time"
col_depth    <- "Depth"
col_measure  <- "DO_conc"

# Plot Labels
lab_datetime <- "Date Time"
lab_depth    <- "Depth (m)"
lab_measure  <- "Dissolved Oxygen (mg/L)"
lab_title    <- "Test Lake"

# Create Plot
p_profile2 <- plot_depth(data = data
                        , col_datetime = col_datetime
                        , col_depth = col_depth
                        , col_measure = col_measure
                        , lab_datetime = lab_datetime
                        , lab_depth = lab_depth
                        , lab_measure = lab_measure
                        , lab_title = lab_title)

# Print Plot
print(p_profile2)
```


## plot_heatmap

```{r plot_heatmap}
library(LakeMonitoR)

# Data (Test Lake)
data         <- laketest

# Column Names
col_datetime <- "Date.Time"
col_depth    <- "Depth"
col_measure  <- "temp_F"

# Plot Labels
lab_datetime <- "Date Time"
lab_depth    <- "Depth (m)"
lab_measure  <- "Temperature (F)"
lab_title    <- "Test Lake"
line_val     <- 2

# Create Plot
p_hm <- plot_heatmap(data = data
                        , col_datetime = col_datetime
                        , col_depth = col_depth
                        , col_measure = col_measure
                        , lab_datetime = lab_datetime
                        , lab_depth = lab_depth
                        , lab_measure = lab_measure
                        , lab_title = lab_title
                        , contours = TRUE)

# Print Plot
print(p_hm)
```

## plot_ts

```{r plot_ts}
# Data (Test Lake)
data         <- laketest_wind

# Column Names
col_datetime <- "Date.Time"
col_measure  <- "WSPD"

# Plot Labels
lab_datetime <- "Date Time"
lab_measure  <- "Average Wind Speed (m/s)"
lab_title    <- "Test Lake"

# Create Plot
p_ts <- plot_ts(data = data
                , col_datetime = col_datetime
                , col_measure = col_measure
                , lab_datetime = lab_datetime
                , lab_measure = lab_measure
                , lab_title = lab_title)

# Print Plot
print(p_ts)
```

## plot_oxythermal
For oxythermal habitat for coldwater fish a plot that shows depths at which 
suitable oxythermal habitat exists during deployment period.  The user can 
enter thresholds for temperature and dissolved oxygen.

```{r plot_oxythermal}
# Data (Test Lake)
data         <- laketest

# Column Names
col_datetime <- "Date.Time"
col_depth    <- "Depth"
col_temp     <- "temp_F"
col_do       <- "DO_conc"

# Data values
thresh_temp <- 68
operator_temp <- "<="
thresh_do <- 3
operator_do <- ">="

# Plot Labels
lab_datetime <- "Date Time"
lab_depth    <- "Depth (m)"
lab_temp     <- "Temp (F)"
lab_title    <- "Test Lake"

# Create Plot
p_ot <- plot_oxythermal(data = data
                        , col_datetime = col_datetime
                        , col_depth = col_depth
                        , col_temp = col_temp
                        , col_do = col_do
                        , thresh_temp = thresh_temp
                        , operator_temp= operator_temp
                        , thresh_do = thresh_do
                        , operator_do = operator_do
                        , lab_datetime = lab_datetime
                        , lab_depth = lab_depth
                        , lab_temp = lab_temp
                        , lab_title = lab_title)

## Add Subtitle and Caption
myST <- paste0("Temp ", operator_temp, " ", thresh_temp
             , " and DO ", operator_do, " ", thresh_do)
p_ot <- p_ot + 
  ggplot2::labs(subtitle = myST) + 
  ggplot2::labs(caption = paste0("Gray = Areas outside of given temp and DO "
                , "values."))

print(p_ot)

# For Comparison, heatmaps of Temperature and DO

# heat map, Temp
p_hm_temp <- plot_heatmap(data = data
                    , col_datetime = col_datetime
                    , col_depth = col_depth
                    , col_measure = col_temp
                    , lab_datetime = lab_datetime
                    , lab_depth = lab_depth
                    , lab_measure = lab_temp
                    , lab_title = lab_title
                    , contours = FALSE)
print(p_hm_temp)

# heat map, DO
p_hm_do <- plot_heatmap(data = data
                         , col_datetime = col_datetime
                         , col_depth = col_depth
                         , col_measure = col_do
                         , lab_datetime = lab_datetime
                         , lab_depth = lab_depth
                         , lab_measure = "DO (mg/L)"
                         , lab_title = lab_title
                         , contours = FALSE)
print(p_hm_do)

# Plot, Combine all 3
gridExtra::grid.arrange(p_ot, p_hm_temp, p_hm_do)
```

## Combining Plots
The shiny app combines the depth plot with the heat map and saves it to the 
Results folder.  Combining the plots is not a default feature of the package
but can be done with the use of the `gridExtra` package.

Move legends to the bottom and remove titles to get more space.

```{r plot_combine}
library(LakeMonitoR)

# Data (Test Lake)
data         <- laketest

# depth----
# Column Names
col_datetime <- "Date.Time"
col_depth    <- "Depth"
col_measure  <- "DO_conc"

# Plot Labels
lab_datetime <- "Date Time"
lab_depth    <- "Depth (m)"
lab_measure  <- "Dissolved Oxygen (mg/L)"
lab_title    <- "Test Lake"

# Create Plot
p_profile <- plot_depth(data = data
                        , col_datetime = col_datetime
                        , col_depth = col_depth
                        , col_measure = col_measure
                        , lab_datetime = lab_datetime
                        , lab_depth = lab_depth
                        , lab_measure = lab_measure
                        , lab_title = NA)

# heat map----
# Data (Test Lake)
data         <- laketest

# Column Names
col_datetime <- "Date.Time"
col_depth    <- "Depth"
col_measure  <- "temp_F"

# Plot Labels
lab_datetime <- "Date Time"
lab_depth    <- "Depth (m)"
lab_measure  <- "Temperature (F)"
lab_title    <- "Test Lake"
line_val     <- 2

# Create Plot----
p_hm <- plot_heatmap(data = data
                        , col_datetime = col_datetime
                        , col_depth = col_depth
                        , col_measure = col_measure
                        , lab_datetime = lab_datetime
                        , lab_depth = lab_depth
                        , lab_measure = lab_measure
                        , lab_title = NA
                        , contours = TRUE)

# Change legend location
p_profile <- p_profile + ggplot2::theme(legend.position = "bottom") +
  ggplot2::geom_hline(yintercept = 2
                      , color = "red"
                      , linetype = "dashed"
                      , size = 1.5)
p_hm <- p_hm + ggplot2::theme(legend.position = "bottom")

# # plot, Wind
p_ts <- plot_ts(data = laketest_wind
                , col_datetime = col_datetime
                , col_measure = "WSPD"
                , lab_datetime = lab_datetime
                , lab_measure = "Average Wind (m/s)"
                , lab_title = NA)

# Combine 2
p_combo_2 <- gridExtra::grid.arrange(p_hm, p_profile)

# Combine 3
p_combo_3 <- gridExtra::grid.arrange(p_hm, p_profile, p_ts)

```

## stratification
Stratification is calculated as greater than or equal to a 1 degree C difference
over 1-m.  **Unless** there are only 2 measurements then stratification is 
calculated as greater than or equal to a 1 degree C difference (i.e., depth is 
not considered).

The example below calculates stratification on the entire data set and then 
again with only the top and bottom measurements.  Using only the top and bottom
measurements results in more days of stratification and has 1 more event.

```{r stratification}
library(LakeMonitoR)

## Example 1, entire data set

# data
data <- laketemp_ddm

# Columns
col_date    <- "Date"
col_depth   <- "Depth"
col_measure <- "Measurement"

# Calculate Stratification
ls_strat <- stratification(data
                           , col_date
                           , col_depth
                           , col_measure
                           , min_days = 1 )

# Results, Stratification Dates
head(ls_strat$Stratification_Dates)

# Results, Stratification Events
ls_strat$Stratification_Events

# Results, Stratification Plot
p_strat <- ls_strat$Stratification_Plot
p_strat <- p_strat + ggplot2::labs(caption = paste0("Stratification >= "
                                           , "1 deg C difference over 1-m"))
print(p_strat)

#~~~~~~~~~~~~

## Example 2, only top and bottom measurement

# data
data <- laketemp_ddm
min_depth <- min(data[, col_depth], na.rm = TRUE)
max_depth <- max(data[, col_depth], na.rm = TRUE)
data_tb <- data[data[, col_depth] == min_depth |
             data[, col_depth] == max_depth, ]

# Columns
col_date    <- "Date"
col_depth   <- "Depth"
col_measure <- "Measurement"

# Calculate Stratification
ls_strat_tb <- stratification(data_tb
                             , col_date
                             , col_depth
                             , col_measure
                             , min_days = 1 )
# Results, Stratification Dates
head(ls_strat_tb$Stratification_Dates)

# Results, Stratification Events
ls_strat_tb$Stratification_Events

# Results, Stratification Plot
p_strat_tb <- ls_strat_tb$Stratification_Plot
p_strat_tb <- p_strat_tb + ggplot2::labs(caption = paste0("Stratification >="
                                          , " 1 deg C (only 2 depths)"))
print(p_strat_tb)
```
## TDOx
Temperature at Dissolved Oxygen x value.

```{r tdox}
# data
data <- laketest

# Columns
col_date  <- "Date.Time"
col_depth <- "Depth"
col_temp  <- "temp_F"
col_do    <- "DO_conc"
do_x_val  <- 3

tdox_3 <- tdox(data = data
               , col_date = col_date
               , col_depth = col_depth
               , col_temp = col_temp
               , col_do = col_do
               , do_x_val = do_x_val)

head(tdox_3)
```

Example plot of a single depth using base R graphics.

Temperature is in black and Dissolved Oxygen is in blue.

Dotted lines are DO = 3 and Temperature = 53.7 (TDO3 value).

```{r plot_TDO3, fig.show='asis'}
# Data
data <- laketest

# Columns
col_date  <- "Date.Time"
col_depth <- "Depth"
col_temp  <- "temp_F"
col_do    <- "DO_conc"
do_x_val  <- 3

df_plot <- data[data[, "Date.Time"] == "2018-07-03 12:00", ]

tdox_3_plot <- tdox(data = df_plot
                    , col_date = col_date
                    , col_depth = col_depth
                    , col_temp = col_temp
                    , col_do = col_do
                    , do_x_val = do_x_val)

knitr::kable(tdox_3_plot)

# Plot
## Temp, x-bottom, circles
plot(df_plot$temp_F, df_plot$Depth, col = "black"
     , ylim = rev(range(df_plot$Depth))
     , xlab = "Temperature (F)", ylab = "Depth (m)")
lines(df_plot$temp_F, df_plot$Depth, col = "black")
abline(v = 53.66, col = "black", lty = 3, lwd = 2)
## plot on top of existing plot
par(new=TRUE)
## Add DO
plot(df_plot$DO_conc, df_plot$Depth, col = "blue", pch = 17
     , ylim = rev(range(df_plot$Depth))
     , xaxt = "n", yaxt = "n", xlab = NA, ylab = NA)
lines(df_plot$DO_conc, df_plot$Depth, col = "blue")
axis(side = 3, col.axis = "blue", col.lab = "blue")
mtext("Dissolved Oxygen (mg/L)", side = 3, line = 3, col = "blue")
abline(v = 3, col = "blue", lty = 3, lwd = 2)
## Legend
legend("bottomright"
      , c("Temperature", "Temp = 53.66 @ DO = 3", "Dissolved Oxygen", "DO = 3")
      , col = c("black", "black", "blue", "blue")
      , pch = c(21, NA, 17, NA)
      , lty = c(1, 3, 1, 3)
      , lwd = c(1, 2, 1, 2))
```


## export_rLakeAnalyzer
Reorganize file for use with `rLakeAnalyzer` package.

```{r}
library(LakeMonitoR)

# Convert Data for use with rLakeAnalyzer

# Data
data <- laketemp_ddm

# Columns, date listed first
col_depth <- "Depth"
col_data <- c("Date", "Measurement")
col_rLA  <- c("datetime", "wtr")

# Run function
df_rLA <- export_rLakeAnalyzer(data, col_depth, col_data, col_rLA)

#
str(df_rLA)

## Not run: 
# Save
#write.csv(df_rLA, file.path(tempdir(), "example_rLA.csv"), row.names = FALSE)
```

# rLakeAnalyzer
After exporting file format for `rLakeAnalyzer` can use the functions from that 
package.

## Help

```{r help, echo = FALSE}
library(rLakeAnalyzer)
print(lsf.str("package:rLakeAnalyzer"))
```

## approx.bathy
Estimate hypsography curve.

```{r approx.bathy}
Voldev.ex <- approx.bathy(Zmax = 25, Zmean = 12, lkeArea = 39400000, method = "voldev")
Voldevshallow.ex <- approx.bathy(Zmax = 25, Zmean = 6, lkeArea = 39400000, method = "voldev")
Cone.ex <- approx.bathy(Zmax = 25, lkeArea = 39400000, method = "cone")

# plot depth-area curves
plot(Cone.ex$depths ~ Cone.ex$Area.at.z, xlab = "Area (m^3)", ylab = "Depth (m)")
points(Voldev.ex$depths ~ Voldev.ex$Area.at.z, col = "red")
points(Voldevshallow.ex$depths ~ Voldevshallow.ex$Area.at.z, col = "blue")
```

## buoyancy.freq
Calculate buoyancy frequency.

```{r buoyancy.freq}
# A vector of water temperatures
wtr <- c(22.51, 22.42, 22.4, 22.4, 22.4, 22.36, 22.3, 22.21, 22.11, 21.23, 16.42, 
	15.15, 14.24, 13.35, 10.94, 10.43, 10.36, 9.94, 9.45, 9.1, 8.91, 8.58, 8.43)

#A vector defining the depths
depths <- c(0, 0.5, 1, 1.5, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 
	17, 18, 19, 20)

b.f <- buoyancy.freq(wtr, depths)

plot(b.f, attr(b.f, 'depths'), type='b'
     , ylab='Depth', xlab='Buoyancy Frequency'
     , ylim=c(max(depths), min(depths)))
```

## center.buoyancy
Calculate the center of buoyancy.

```{r center.buoyancy}
# A vector of water temperatures
wtr <- c(22.51, 22.42, 22.4, 22.4, 22.4, 22.36, 22.3, 22.21, 22.11, 21.23, 16.42, 
	15.15, 14.24, 13.35, 10.94, 10.43, 10.36, 9.94, 9.45, 9.1, 8.91, 8.58, 8.43)

#A vector defining the depths
depths <- c(0, 0.5, 1, 1.5, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 
	17, 18, 19, 20)

c.b <- center.buoyancy(wtr, depths)
c.b
```

## epi.temperature
Get volumetrically averaged epilimnion temp.

```{r epi.temperature}
## No Example
```

## hypo.temperature
Get volumetrically averaged hypolimnion temp.

```{r hypo.temperature}
##No Example
```

## internal.energy
Internal energy function (Joules).

```{r internal.energy}
bthA   <-	c(1000,900,864,820,200,10)
bthD	 <-	c(0,2.3,2.5,4.2,5.8,7)

wtr	   <-	c(28,27,26.4,26,25.4,24,23.3)
depths <-	c(0,1,2,3,4,5,6)

cat('Internal Energy for input is: ')
cat(internal.energy(wtr, depths, bthA, bthD))
```

## lake.number
Calculate Lake Number.

```{r lake.number}
bthA	<-	c(1000,900,864,820,200,10)
bthD	<-	c(0,2.3,2.5,4.2,5.8,7)
uStar	<-	c(0.0032,0.0024)
St	  <-	c(140,153)
metaT	<-	c(1.34,1.54)
metaB	<-	c(4.32,4.33)
averageHypoDense	<-	c(999.3,999.32)
cat('Lake Number for input vector is: ')
cat(lake.number( bthA, bthD, uStar, St, metaT, metaB, averageHypoDense) )
```

## layer.density
Returns the average density of a layer between two depths.

```{r layer.density}
top     <- 2
bottom  <- 6
wtr     <- c(25.2,25.1,24.1,22.0,19.8,15.3,12.0,11.1)
depths  <- c(0,1,2,3,4,5,6,7) 
bthA    <- c(10000,8900,5000,3500,2000,1000,300,10)
bthD    <- c(0,1,2,3,4,5,6,7)
layer.density(top,bottom,wtr,depths,bthA,bthD)
```

## layer.temperature
Returns the average temperature of a layer between two depths.

```{r layer.temperature}
# Supply input data
top     <- 2
bottom  <- 6
wtr     <- c(25.2,25.1,24.1,22.0,19.8,15.3,12.0,11.1)
depths  <- c(0,1,2,3,4,5,6,7) 
bthA    <- c(10000,8900,5000,3500,2000,1000,300,10)
bthD    <- c(0,1,2,3,4,5,6,7)

#Return the average temperature of the water column between 2 and 6 meters.
layer.temperature(top,bottom,wtr,depths,bthA,bthD)
```

## meta.depths
Calculate the Top and Bottom Depths of the Metalimnion.

To get the thickness of Metalimnion subtract the two values.

```{r meta.depths}
wtr = c(22.51, 22.42, 22.4, 22.4, 22.4, 22.36, 22.3, 22.21, 22.11, 21.23, 16.42, 
15.15, 14.24, 13.35, 10.94, 10.43, 10.36, 9.94, 9.45, 9.1, 8.91, 8.58, 8.43)

depths = c(0, 0.5, 1, 1.5, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 
			 17, 18, 19, 20)
			 
m.d = meta.depths(wtr, depths, slope=0.1, seasonal=FALSE)

cat('The top depth of the metalimnion is:', m.d[1])

cat('The bottom depth of the metalimnion is:', m.d[2])

cat("The thickness of the metalimnion is: ", m.d[2] - m.d[1])

```

## schmidt.stability
Calculate the Schmidt stability.

```{r schmidt.stability}
bthA    <- c(1000,900,864,820,200,10)
bthD    <- c(0,2.3,2.5,4.2,5.8,7)

wtr	    <- c(28,27,26.4,26,25.4,24,23.3)
depths	<- c(0,1,2,3,4,5,6)

cat('Schmidt stability for input is: ')
cat(schmidt.stability(wtr, depths, bthA, bthD))
```

## thermo.depth
Calculate depth of the thermocline from a temperature profile

```{r thermo.depth}
# A vector of water temperatures
wtr <- c(22.51, 22.42, 22.4, 22.4, 22.4, 22.36, 22.3, 22.21, 22.11, 21.23, 16.42, 
		15.15, 14.24, 13.35, 10.94, 10.43, 10.36, 9.94, 9.45, 9.1, 8.91, 8.58, 8.43)

#A vector defining the depths
depths <- c(0, 0.5, 1, 1.5, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 
   17, 18, 19, 20)

t.d <- thermo.depth(wtr, depths, seasonal=FALSE)

cat('The thermocline depth is:', t.d)
```

## ts.buoyancy.freq
Calculate the buoyancy (Brunt-Vaisala) frequency for a temperature profile.

```{r ts.buoyancy.freq}
#Get the path for the package example file included
wtr.path <- system.file('extdata'
                        , 'Sparkling.daily.wtr'
                        , package="rLakeAnalyzer")

#Load data for example lake, Sparkilng Lake, Wisconsin.
sp.wtr <- load.ts(wtr.path)

N2  <- ts.buoyancy.freq(sp.wtr, seasonal=FALSE)
SN2 <- ts.buoyancy.freq(sp.wtr, seasonal=TRUE)

plot(N2, type='l', ylab='Buoyancy Frequency', xlab='Date')
lines(SN2, col='red')
```

### ts.buoyancy.freq, LakeMonitoR
Using LakeMonitoR data.

```{r}
library(LakeMonitoR)

# Convert Data for use with rLakeAnalyzer

# Data
data <- laketemp_ddm

# Columns, date listed first
col_depth <- "Depth"
col_data <- c("Date", "Measurement")
col_rLA  <- c("datetime", "wtr")

# Run function
df_rLA <- export_rLakeAnalyzer(data, col_depth, col_data, col_rLA)

# Calculate
N2  <- ts.buoyancy.freq(df_rLA, seasonal=FALSE)
SN2 <- ts.buoyancy.freq(df_rLA, seasonal=TRUE)

knitr::kable(head(N2))
knitr::kable(head(SN2))

plot(N2, type='l', ylab='Buoyancy Frequency', xlab='Date', lwd = 3)
lines(SN2, col='red')
legend("topleft"
       , legend = c("N2", "SN2")
       , col = c("black", "red")
       , lwd = c(1, 3))
```

## ts.center.buoyancy
Calculates the center of buoyancy for multiple temperature profiles.

```{r ts.center.buoyancy}
#Get the path for the package example file included
wtr.path <- system.file('extdata', 'Sparkling.daily.wtr', package="rLakeAnalyzer")

#Load data for example lake, Sparkilng Lake, Wisconsin.
sp.wtr <- load.ts(wtr.path)
 
#calculate and plot the thermocline depth
t.d <- ts.thermo.depth(sp.wtr)

center.N2 = ts.center.buoyancy(sp.wtr)

plot(center.N2, type='l', ylab='Depth (m)', xlab='Date', ylim=c(19,0), lwd = 1.5)
lines(t.d, type='l', col='red', lwd = 1.5)
legend(x = t.d[3,1], y = .25,
   c('center of buoyancy','thermocline depth'),
   lty=c(1,1),
   lwd=c(1.5,1.5),col=c("black","red"), bty = "n")
```

### ts.center.buoyancy, LakeMonitoR
Using LakeMonitoR data.

```{r}
library(LakeMonitoR)

# Convert Data for use with rLakeAnalyzer

# Data
data <- laketemp_ddm

# Columns, date listed first
col_depth <- "Depth"
col_data <- c("Date", "Measurement")
col_rLA  <- c("datetime", "wtr")

# Run function
df_rLA <- export_rLakeAnalyzer(data, col_depth, col_data, col_rLA)

# Calculate
t.d <- ts.thermo.depth(df_rLA)

center.N2 = ts.center.buoyancy(df_rLA)

knitr::kable(head(center.N2))

plot(center.N2, type='l', ylab='Depth (m)', xlab='Date', ylim=c(19,0), lwd = 1.5)
lines(t.d, type='l', col='red', lwd = 1.5)
legend(x = t.d[3,1], y = .25,
   c('center of buoyancy','thermocline depth'),
   lty=c(1,1),
   lwd=c(1.5,1.5),col=c("black","red"), bty = "n")
```

## ts.internal.energy
Calculate physical indices for a timeseries.

```{r ts.internal.energy}
# No Example
```

## ts.lake.number
Calculate physical indices for a timeseries.

```{r ts.lake.number}
# No Example
```

## ts.layer.temperature
Calculate volume-weighted average water temperature across a range of depths for a timeseries.

```{r ts.layer.temperature}
#Get the path for the package example file included
wtr.path   <- system.file('extdata', 'Sparkling.daily.wtr', package="rLakeAnalyzer")
bathy.path <- system.file('extdata', 'Sparkling.bth', package="rLakeAnalyzer")

#Load data for example lake, Sparkilng lake, in Wisconsin.
sp.wtr   <- load.ts(wtr.path)
sp.bathy <- load.bathy(bathy.path)


l.t <- ts.layer.temperature(sp.wtr, 0, 18, sp.bathy)
plot(l.t$datetime, l.t$layer.temp, type='l', 
      ylab='Volumetrically averaged lake temperature', xlab='Date')

# ggplot
ggplot2::ggplot(data = l.t, ggplot2::aes(x = datetime, y = layer.temp)) +
  ggplot2::geom_line(na.rm = TRUE) + 
  ggplot2::labs(x = "Date", y = "Volumetrically averaged lake temperature") +
  ggplot2::theme_bw()
```

## ts.meta.depths
Calculate physical indices for a timeseries.

```{r ts.meta.depths, }
#Get the path for the package example file included
exampleFilePath <- system.file('extdata'
                               , 'Sparkling.daily.wtr'
                               , package="rLakeAnalyzer")

#Load
sparkling.temp <- load.ts(exampleFilePath)

# Calculate and plot the metalimnion depths
m.d <- ts.meta.depths(sparkling.temp)
m.d$thickness <- m.d$bottom - m.d$top

knitr::kable(m.d[1:10, ])

plot(m.d$datetime
     , m.d$top
     , type='l'
     , ylab='Meta Depths (m)'
     , xlab='Date'
     , col='blue')
lines(m.d$datetime, m.d$bottom, col='red')

# ggplot
ggplot2::ggplot(data = m.d, ggplot2::aes(x = datetime, y = top)) +
  ggplot2::geom_line(color = "blue", na.rm = TRUE) + 
  ggplot2::labs(x = "Date", y = "Meta Depths (m)") +
  ggplot2::theme_bw() + 
  ggplot2::geom_line(data = m.d
                     , ggplot2::aes(x = datetime, y = bottom)
                     , color = "red"
                     , na.rm = TRUE)
```

### ts.meta.depths, LakeMonitoR
Using LakeMonitoR data.

Then calculate for epilimnion and hypolimnion temp difference.

Red is hypolimnion and blue is metalimnion.

```{r}
library(LakeMonitoR)
library(dplyr)
library(tidyr)

# Convert Data for use with rLakeAnalyzer

# Data
data <- laketemp_ddm

# Columns, date listed first
col_depth <- "Depth"
col_data <- c("Date", "Measurement")
col_rLA  <- c("datetime", "wtr")

# Run function
df_rLA <- export_rLakeAnalyzer(data, col_depth, col_data, col_rLA)

# Calculate
m.d <- ts.meta.depths(df_rLA)
m.d$thickness <- m.d$bottom - m.d$top

knitr::kable(m.d[1:10, ])

# ggplot
ggplot2::ggplot(data = m.d, ggplot2::aes(x = datetime, y = top)) +
  ggplot2::geom_line(color = "blue", na.rm = TRUE) + 
  ggplot2::labs(x = "Date", y = "Meta Depths (m)") +
  ggplot2::theme_bw() + 
  ggplot2::geom_line(data = m.d
                     , ggplot2::aes(x = datetime, y = bottom)
                     , color = "red"
                     , na.rm = TRUE)


# temp diff epi and hypo
df_merge <- merge(data, m.d, by.x = "Date", by.y = "datetime")

df_merge[, "Epi"] <- ifelse(df_merge[, "Depth"] < df_merge[,  "top"], TRUE, FALSE)
df_merge[, "Meta"] <- NA
df_merge[, "Hypo"] <- ifelse(df_merge[, "Depth"] > df_merge[,  "bottom"], TRUE, FALSE)
df_merge[, "Meta"] <- ifelse(df_merge[, "Epi"] + df_merge[, "Hypo"] == 0, TRUE, FALSE)
df_merge[, "Layer"] <- ifelse(df_merge[, "Epi"] == TRUE
                            , "Epi"
                            , ifelse(df_merge[, "Meta"] ==  TRUE
                                  , "Meta"
                                  , ifelse(df_merge[, "Hypo"] ==  TRUE
                                          , "Hypo", NA)))

col_datetime <- "Date"
col_depth <- "Depth"
col_measure <- "Measurement"
col_layer <- "Layer"

df_merge[, "Layer"] <- factor(df_merge[, "Layer"]
                               , levels = c("Epi", "Meta", "Hypo", NA))

stats_layers <- df_merge %>%
      dplyr::group_by(.data[[col_datetime]], .data[[col_layer]]) %>%
      dplyr::summarize(.groups = "keep"
                       #, groupname = col_datetime
                       , n = length(.data[[col_measure]])
                       , ndays = length(unique(.data[[col_datetime]]))
                       , mean = mean(.data[[col_measure]], na.rm = TRUE)
                       , median = stats::median(.data[[col_measure]]
                                                , na.rm = TRUE)
                       , min = min(.data[[col_measure]], na.rm = TRUE)
                       , max = max(.data[[col_measure]], na.rm = TRUE)
                       , range = max - min
                       , sd = stats::sd(.data[[col_measure]], na.rm = TRUE)
                       , var = stats::var(.data[[col_measure]], na.rm = TRUE)
                       , cv = sd/mean
                       , q01 = stats::quantile(.data[[col_measure]]
                                               , probs = .01
                                               , na.rm = TRUE)
                       , q05 = stats::quantile(.data[[col_measure]]
                                               , probs = .05
                                               , na.rm = TRUE)
                       , q10 = stats::quantile(.data[[col_measure]]
                                               , probs = .10
                                               , na.rm = TRUE)
                       , q25 = stats::quantile(.data[[col_measure]]
                                               , probs = .25
                                               , na.rm = TRUE)
                       , q50 = stats::quantile(.data[[col_measure]]
                                               , probs = .50
                                               , na.rm = TRUE)
                       , q75 = stats::quantile(.data[[col_measure]]
                                               , probs = .75
                                               , na.rm = TRUE)
                       , q90 = stats::quantile(.data[[col_measure]]
                                               , probs = .90
                                               , na.rm = TRUE)
                       , q95 = stats::quantile(.data[[col_measure]]
                                               , probs = .95
                                               , na.rm = TRUE)
                       , q99 = stats::quantile(.data[[col_measure]]
                                               , probs = .99
                                               , na.rm = TRUE)
                       #, n_below = sum(.data[["below"]], na.rm = TRUE)
      )

# pivot wider, min layer temp

layers_min <- pivot_wider(stats_layers
                          , id_cols = "Date"
                          , names_from = "Layer"
                          , names_sort = TRUE
                          , values_from = "min")

layers_min[, "diff_EH"] <- layers_min[, "Hypo"] - layers_min[, "Epi"]

head(layers_min)

plot(layers_min$diff_EH, type = "l")

ggplot2::ggplot(data = layers_min, ggplot2::aes(x = Date, y = diff_EH)) +
  ggplot2::geom_line(color = "black", na.rm = TRUE) + 
  ggplot2::labs(x = "Date"
                , y = "Temperature Difference (C)"
                , title = "Comparison of Epilimnion and Hypolimnion") +
  ggplot2::theme_bw()  


```


## ts.schmidt.stability
Calculate physical indices for a timeseries.

```{r ts.schmidt.stability}
# No Example
```

### ts.schmidt.stability, LakeMonitor
Using LakeMonitoR data.  Requires bathymetry data.

```{r, eval=FALSE}
library(LakeMonitoR)

# Convert Data for use with rLakeAnalyzer

# Data
data <- laketemp_ddm
bathy <- NA

# Columns, date listed first
col_depth <- "Depth"
col_data <- c("Date", "Measurement")
col_rLA  <- c("datetime", "wtr")

# Run function
df_rLA <- export_rLakeAnalyzer(data, col_depth, col_data, col_rLA)


# Calculate
s.s <- ts.schmidt.stability(df_rLA, bathy)

```

## ts.thermo.depth
Calculate physical indices for a timeseries.

```{r ts.thermo.depth}
#Get the path for the package example file included
exampleFilePath <- system.file('extdata', 'Sparkling.daily.wtr', package="rLakeAnalyzer")

#Load
sparkling.temp <- load.ts(exampleFilePath)


#calculate and plot the thermocline depth
t.d <- ts.thermo.depth(sparkling.temp)

plot(t.d$datetime, t.d$thermo.depth, type='l', ylab='Thermocline Depth (m)', xlab='Date')

# ggplot
ggplot2::ggplot(data = t.d, ggplot2::aes(x = datetime, y = thermo.depth)) +
  ggplot2::geom_line(na.rm = TRUE) +
  ggplot2::labs(x = "Date", y = "Thermocline Depth (m)") + 
  ggplot2::theme_bw()
```

### ts.thermo.depth, LakeMonitoR
Using LakeMonitoR data.

```{r}
library(LakeMonitoR)

# Convert Data for use with rLakeAnalyzer

# Data
data <- laketemp_ddm

# Columns, date listed first
col_depth <- "Depth"
col_data <- c("Date", "Measurement")
col_rLA  <- c("datetime", "wtr")

# Run function
df_rLA <- export_rLakeAnalyzer(data, col_depth, col_data, col_rLA)

# Calculate
t.d <- ts.thermo.depth(df_rLA)

knitr::kable(t.d)

plot(t.d$datetime, t.d$thermo.depth, type='l', ylab='Thermocline Depth (m)', xlab='Date')

# ggplot
ggplot2::ggplot(data = t.d, ggplot2::aes(x = datetime, y = thermo.depth)) +
  ggplot2::geom_line(na.rm = TRUE) +
  ggplot2::labs(x = "Date", y = "Thermocline Depth (m)") + 
  ggplot2::theme_bw()
```

## ts.uStar
Calculate physical indices for a timeseries.

```{r ts.uStar}
# No Example
```

## ts.wedderburn.number
Calculate Wedderburn number for a timeseries.

```{r ts.wedderburn.number}
#Get the path for the package example file included
wtr.path   <- system.file('extdata', 'Sparkling.daily.wtr', package="rLakeAnalyzer")
wnd.path   <- system.file('extdata', 'Sparkling.daily.wnd', package="rLakeAnalyzer")
bathy.path <- system.file('extdata', 'Sparkling.bth', package="rLakeAnalyzer")

#Load data for example lake, Sparkilng lake, in Wisconsin.
sp.wtr   <-  load.ts(wtr.path)
sp.wnd   <- load.ts(wnd.path)
sp.bathy <- load.bathy(bathy.path)

sp.area = 64e4  #Area of Sparkling lake in m^2
wnd.height = 2  #Height of Sparkling lake anemometer

w.n <- ts.wedderburn.number(sp.wtr, sp.wnd, wnd.height, sp.bathy, sp.area)
plot(w.n$datetime, w.n$wedderburn.number, type='l', ylab='Wedderburn Number', xlab='Date')

# ggplot
ggplot2::ggplot(data = w.n, ggplot2::aes(x = datetime, y = wedderburn.number)) +
  ggplot2::geom_line(na.rm = TRUE) +
  ggplot2::labs(x = "Date", y = "Wedderburn Number") + 
  ggplot2::theme_bw()
```

## uStar
Calculates the water friction velocity, uStar.

```{r uStar}
wndSpeed  <- c(5.1,6.3,6.3,5.2,7,7.2)
wndHeight	<-	2
averageEpiDense	<- c(14,15,14.2,13,12,12)

cat('uStar for input vector is: ')
cat(uStar(wndSpeed,wndHeight,averageEpiDense))
```

## water.density
Estimate Water Density.

```{r water.density}
#Plot water density for water between 1 and 30 deg C
dens <- water.density(1:30)
plot(1:30, dens, xlab="Temp(deg C)", ylab="Density(kg/m^3)")

# ggplot
ggplot2::ggplot() + 
  ggplot2::geom_point(ggplot2::aes(x = 1:30, y = dens)) + 
  ggplot2::labs(x = "Temp (deg C)", y = "Density (kg/m^3)") + 
  ggplot2::theme_bw()
```

## wedderburn.number
Calculates Wedderburn Number for a lake.

```{r wedderburn.number}
delta_rho <- c(3.1,1.5)
metaT     <- c(5.5,2.4)
uSt       <- c(0.0028,0.0032)
Ao        <- c(80300,120000)
AvHyp_rho <- c(999.31,999.1)
wedderburn.number(delta_rho, metaT, uSt, Ao, AvHyp_rho)
```

## whole.lake.temperature
Get volumetrically averaged whole lake temperature.

```{r whole.lake.temperature}
# No Example
```

## wtr.layer
Exploration of lake water column layers.

```{r wtr.layer}
data("latesummer")
df1 <- wtr.layer(depth=latesummer$depth, measure = latesummer$temper)
df1$mld
df1$segments

wtr.layer(data = latesummer, depth=depth, measure = temper, nseg=4)
```

