---
title: "rLakeAnalyzer Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rLakeAnalyzer Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.dim = c(8, 6)/1.5
  #out.width = "50%"
)
library(LakeMonitoR)
```

# Purpose
Show users how to use functions in rLakeAnalyzer from LakeMonitor.

In most cases are using examples provided by rLakeAnalyzer.

# LakeMonitoR

## agg_depth_files

## daily_depth_means

## lake_summary_stats

## plot_depth

## plot_heatmap

## export_rLakeAnalyzer

# rLakeAnalyzer

## Help
```{r help, echo = FALSE}
library(rLakeAnalyzer)
print(lsf.str("package:rLakeAnalyzer"))
```
## approx.bathy
Estimate hypsography curve

```{r approx.bathy}
Voldev.ex <- approx.bathy(Zmax = 25, Zmean = 12, lkeArea = 39400000, method = "voldev")
Voldevshallow.ex <- approx.bathy(Zmax = 25, Zmean = 6, lkeArea = 39400000, method = "voldev")
Cone.ex <- approx.bathy(Zmax = 25, lkeArea = 39400000, method = "cone")

# plot depth-area curves
plot(Cone.ex$depths ~ Cone.ex$Area.at.z, xlab = "Area (m^3)", ylab = "Depth (m)")
points(Voldev.ex$depths ~ Voldev.ex$Area.at.z, col = "red")
points(Voldevshallow.ex$depths ~ Voldevshallow.ex$Area.at.z, col = "blue")
```


## buoyancy.freq
Calculate buoyancy frequency
```{r buoyancy.freq}
# A vector of water temperatures
wtr <- c(22.51, 22.42, 22.4, 22.4, 22.4, 22.36, 22.3, 22.21, 22.11, 21.23, 16.42, 
	15.15, 14.24, 13.35, 10.94, 10.43, 10.36, 9.94, 9.45, 9.1, 8.91, 8.58, 8.43)

#A vector defining the depths
depths <- c(0, 0.5, 1, 1.5, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 
	17, 18, 19, 20)

b.f <- buoyancy.freq(wtr, depths)

plot(b.f, attr(b.f, 'depths'), type='b'
     , ylab='Depth', xlab='Buoyancy Frequency'
     , ylim=c(max(depths), min(depths)))

```

## center.buoyancy
Calculate the center of buoyancy
```{r center.buoyancy}
# A vector of water temperatures
wtr <- c(22.51, 22.42, 22.4, 22.4, 22.4, 22.36, 22.3, 22.21, 22.11, 21.23, 16.42, 
	15.15, 14.24, 13.35, 10.94, 10.43, 10.36, 9.94, 9.45, 9.1, 8.91, 8.58, 8.43)

#A vector defining the depths
depths <- c(0, 0.5, 1, 1.5, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 
	17, 18, 19, 20)

c.b <- center.buoyancy(wtr, depths)
c.b
```

## epi.temperature
Get volumetrically averaged epilimnion temp
```{r epi.temperature}
## No Example
```

## hypo.temperature
Get volumetrically averaged hypolimnion temp
```{r hypo.temperature}
##No Example
```

## internal.energy
Internal energy function (Joules)

```{r internal.energy}
bthA   <-	c(1000,900,864,820,200,10)
bthD	 <-	c(0,2.3,2.5,4.2,5.8,7)

wtr	   <-	c(28,27,26.4,26,25.4,24,23.3)
depths <-	c(0,1,2,3,4,5,6)

cat('Internal Energy for input is: ')
cat(internal.energy(wtr, depths, bthA, bthD))
```

## lake.number
Calculate Lake Number
```{r lake.number}
bthA	<-	c(1000,900,864,820,200,10)
bthD	<-	c(0,2.3,2.5,4.2,5.8,7)
uStar	<-	c(0.0032,0.0024)
St	  <-	c(140,153)
metaT	<-	c(1.34,1.54)
metaB	<-	c(4.32,4.33)
averageHypoDense	<-	c(999.3,999.32)
cat('Lake Number for input vector is: ')
cat(lake.number( bthA, bthD, uStar, St, metaT, metaB, averageHypoDense) )
```

## layer.density
Returns the average density of a layer between two depths.
```{r layer.density}
top     <- 2
bottom  <- 6
wtr     <- c(25.2,25.1,24.1,22.0,19.8,15.3,12.0,11.1)
depths  <- c(0,1,2,3,4,5,6,7) 
bthA    <- c(10000,8900,5000,3500,2000,1000,300,10)
bthD    <- c(0,1,2,3,4,5,6,7)
layer.density(top,bottom,wtr,depths,bthA,bthD)
```

## layer.temperature
Returns the average temperature of a layer between two depths.
```{r layer.temperature}
# Supply input data
top     <- 2
bottom  <- 6
wtr     <- c(25.2,25.1,24.1,22.0,19.8,15.3,12.0,11.1)
depths  <- c(0,1,2,3,4,5,6,7) 
bthA    <- c(10000,8900,5000,3500,2000,1000,300,10)
bthD    <- c(0,1,2,3,4,5,6,7)

#Return the average temperature of the water column between 2 and 6 meters.
layer.temperature(top,bottom,wtr,depths,bthA,bthD)
```

## meta.depths
Calculate the Top and Bottom Depths of the Metalimnion
```{r meta.depths}
wtr = c(22.51, 22.42, 22.4, 22.4, 22.4, 22.36, 22.3, 22.21, 22.11, 21.23, 16.42, 
15.15, 14.24, 13.35, 10.94, 10.43, 10.36, 9.94, 9.45, 9.1, 8.91, 8.58, 8.43)

depths = c(0, 0.5, 1, 1.5, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 
			 17, 18, 19, 20)
			 
m.d = meta.depths(wtr, depths, slope=0.1, seasonal=FALSE)
cat('The top depth of the metalimnion is:', m.d[1])
cat('The bottom depth of the metalimnion is:', m.d[2])

```

## schmidt.stability
Calculate the Schmidt stability
```{r schmidt.stability}
bthA    <- c(1000,900,864,820,200,10)
bthD    <- c(0,2.3,2.5,4.2,5.8,7)

wtr	    <- c(28,27,26.4,26,25.4,24,23.3)
depths	<- c(0,1,2,3,4,5,6)

cat('Schmidt stability for input is: ')
cat(schmidt.stability(wtr, depths, bthA, bthD))
```

## thermo.depth
Calculate depth of the thermocline from a temperature profile
```{r thermo.depth}
# A vector of water temperatures
wtr <- c(22.51, 22.42, 22.4, 22.4, 22.4, 22.36, 22.3, 22.21, 22.11, 21.23, 16.42, 
		15.15, 14.24, 13.35, 10.94, 10.43, 10.36, 9.94, 9.45, 9.1, 8.91, 8.58, 8.43)

#A vector defining the depths
depths <- c(0, 0.5, 1, 1.5, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 
   17, 18, 19, 20)

t.d <- thermo.depth(wtr, depths, seasonal=FALSE)

cat('The thermocline depth is:', t.d)
```

## ts.buoyancy.freq
Calculate the buoyancy (Brunt-Vaisala) frequency for a temperature profile
```{r ts.buoyancy.freq}
#Get the path for the package example file included
wtr.path <- system.file('extdata', 'Sparkling.daily.wtr', package="rLakeAnalyzer")

#Load data for example lake, Sparkilng Lake, Wisconsin.
sp.wtr <- load.ts(wtr.path)

N2  <- ts.buoyancy.freq(sp.wtr, seasonal=FALSE)
SN2 <- ts.buoyancy.freq(sp.wtr, seasonal=TRUE)

plot(N2, type='l', ylab='Buoyancy Frequency', xlab='Date')
lines(SN2, col='red')
```

## ts.center.buoyancy
Calculates the center of buoyancy for multiple temperature profiles
```{r ts.center.buoyancy}
#Get the path for the package example file included
wtr.path <- system.file('extdata', 'Sparkling.daily.wtr', package="rLakeAnalyzer")

#Load data for example lake, Sparkilng Lake, Wisconsin.
sp.wtr <- load.ts(wtr.path)
 
#calculate and plot the thermocline depth
t.d <- ts.thermo.depth(sp.wtr)

center.N2 = ts.center.buoyancy(sp.wtr)

plot(center.N2, type='l', ylab='Depth (m)', xlab='Date', ylim=c(19,0), lwd = 1.5)
lines(t.d, type='l', col='red', lwd = 1.5)
legend(x = t.d[3,1], y = .25,
   c('center of buoyancy','thermocline depth'),
   lty=c(1,1),
   lwd=c(1.5,1.5),col=c("black","red"), bty = "n")
```

## ts.internal.energy
Calculate physical indices for a timeseries
```{r ts.internal.energy}
# No Example
```

## ts.lake.number
Calculate physical indices for a timeseries
```{r ts.lake.number}
# No Example
```

## ts.layer.temperature
Calculate volume-weighted average water temperature across a range of depths for a timeseries
```{r ts.layer.temperature}
#Get the path for the package example file included
wtr.path   <- system.file('extdata', 'Sparkling.daily.wtr', package="rLakeAnalyzer")
bathy.path <- system.file('extdata', 'Sparkling.bth', package="rLakeAnalyzer")

#Load data for example lake, Sparkilng lake, in Wisconsin.
sp.wtr   <- load.ts(wtr.path)
sp.bathy <- load.bathy(bathy.path)


l.t <- ts.layer.temperature(sp.wtr, 0, 18, sp.bathy)
plot(l.t$datetime, l.t$layer.temp, type='l', 
      ylab='Volumetrically averaged lake temperature', xlab='Date')
```

## ts.meta.depths
Calculate physical indices for a timeseries
```{r ts.meta.depths}
#Get the path for the package example file included
exampleFilePath <- system.file('extdata', 'Sparkling.daily.wtr', package="rLakeAnalyzer")

#Load
sparkling.temp <- load.ts(exampleFilePath)

#calculate and plot the metalimnion depths
m.d <- ts.meta.depths(sparkling.temp)

plot(m.d$datetime, m.d$top, type='l', ylab='Meta Depths (m)', xlab='Date', col='blue')
lines(m.d$datetime, m.d$bottom, col='red')
```

## ts.schmidt.stability
Calculate physical indices for a timeseries
```{r ts.schmidt.stability}
# No Example
```

## ts.thermo.depth
Calculate physical indices for a timeseries
```{r ts.thermo.depth}
#Get the path for the package example file included
exampleFilePath <- system.file('extdata', 'Sparkling.daily.wtr', package="rLakeAnalyzer")

#Load
sparkling.temp <- load.ts(exampleFilePath)


#calculate and plot the thermocline depth
t.d <- ts.thermo.depth(sparkling.temp)

plot(t.d$datetime, t.d$thermo.depth, type='l', ylab='Thermocline Depth (m)', xlab='Date')
```

## ts.uStar
Calculate physical indices for a timeseries
```{r ts.uStar}
# No Example
```

## ts.wedderburn.number
Calculate Wedderburn number for a timeseries
```{r ts.wedderburn.number}
#Get the path for the package example file included
wtr.path   <- system.file('extdata', 'Sparkling.daily.wtr', package="rLakeAnalyzer")
wnd.path   <- system.file('extdata', 'Sparkling.daily.wnd', package="rLakeAnalyzer")
bathy.path <- system.file('extdata', 'Sparkling.bth', package="rLakeAnalyzer")

#Load data for example lake, Sparkilng lake, in Wisconsin.
sp.wtr   <-  load.ts(wtr.path)
sp.wnd   <- load.ts(wnd.path)
sp.bathy <- load.bathy(bathy.path)

sp.area = 64e4  #Area of Sparkling lake in m^2
wnd.height = 2  #Height of Sparkling lake anemometer

w.n <- ts.wedderburn.number(sp.wtr, sp.wnd, wnd.height, sp.bathy, sp.area)
plot(w.n$datetime, w.n$wedderburn.number, type='l', ylab='Wedderburn Number', xlab='Date')
```

## uStar
Calculates the water friction velocity, uStar
```{r uStar}
wndSpeed  <- c(5.1,6.3,6.3,5.2,7,7.2)
wndHeight	<-	2
averageEpiDense	<- c(14,15,14.2,13,12,12)

cat('uStar for input vector is: ')
cat(uStar(wndSpeed,wndHeight,averageEpiDense))
```

## water.density
Estimate Water Density
```{r water.density}
#Plot water density for water between 1 and 30 deg C
dens <- water.density(1:30)
plot(1:30, dens, xlab="Temp(deg C)", ylab="Density(kg/m^3)")
```

## wedderburn.number
Calculates Wedderburn Number for a lake
```{r wedderburn.number}
delta_rho <- c(3.1,1.5)
metaT     <- c(5.5,2.4)
uSt       <- c(0.0028,0.0032)
Ao        <- c(80300,120000)
AvHyp_rho <- c(999.31,999.1)
wedderburn.number(delta_rho, metaT, uSt, Ao, AvHyp_rho)
```

## whole.lake.temperature
Get volumetrically averaged whole lake temperature
```{r whole.lake.temperature}
# No Example
```

## wtr.layer
Exploration of lake water column layers
```{r wtr.layer}
data("latesummer")
df1 <- wtr.layer(depth=latesummer$depth, measure = latesummer$temper)
df1$mld
df1$segments

wtr.layer(data = latesummer, depth=depth, measure = temper, nseg=4)
```

