---
title: "Notebook, baytrendsmap"
output: html_notebook
---

# Package

Package related code.

## Vignette
```{r, eval=FALSE}
# generate Vignette
library(LakeMonitoR)
library(devtools)
devtools::build_vignettes()
```
devtools::build() more useful and build_vignettes()

## Create Package
Use the code below to create the package.  Assumes you are in the development package directory

```{r, eval=FALSE}
# NEWS
# Render then Copy NEWS so picked up in help
rmarkdown::render("NEWS.rmd", "all")
file.copy("NEWS.md", "NEWS", overwrite = TRUE)
file.remove("NEWS.html")
#file.remove("NEWS.md")
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Library Name
myPkg <- "LakeMonitoR"
# Load Library
library(devtools)
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Create Package
# create(myLibrary)
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# files to library folder
# file.copy("metric.values.R","./Library/MMIcalc/R/metric.values.R",overwrite=TRUE)
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Document, Install, and Reload Library
## Generate Documentation
setwd(paste0("./", myPkg))
devtools::document()
## Install New Package (locally)
setwd("..") # return to root directory first
devtools::install(myPkg, quick=FALSE, reload=TRUE, build_vignettes = TRUE)

## Reload library
library(myPkg, character.only = TRUE)
# change wd back to package
setwd(paste0("./", myPkg))
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

After creating the package reload it after restarting R within RStudio (Control+Shift+F10).

```{r, eval=FALSE}
# Restart R within RStudio:  Ctrl + Shift + F10
myPkg <- "LakeMonitoR"
library(myPkg, character.only = TRUE)
help(package=(myPkg))

#?metric.scores 
citation(myPkg)

# Create PkgDown Website
usethis::use_pkgdown()
pkgdown::build_site()

```

Reinstall from GitHub before upload Shiny app.
```{r, eval=FALSE}
library(devtools)
install_github("leppott/LakeMonitoR", force = TRUE, build_vignettes = TRUE)
```


### Data Documentation
Use promptData to get data.r listing for columns.  Short cuts creating documentation.

```{r DocData}
promptData(laketemp)
```

**code coverage**
```{r CodeCoverage}
#library(covr)
covr::report()
```


## Error Checking
Need to use testthat but not to that stage yet.
```{r, eval=FALSE}
# Check for errors (or press Cmd + Shift + E in RStudio)
#http://r-pkgs.had.co.nz/check.html
devtools::check()
```

# CI, Badges, Code Coverage, etc
* Continuous Integration with GitHub Actions (TravisCI shut down Dec 2020)
* Code Quality (CodeFactor)
    + CodeFactor.io
        - GitHub login
* Code Coverage (testing)
    + Codecov.io
        - GitHub login
        - After set up CI with GitHub Actions
* Package website (pkgdown and GitHub Action)
    + branch gh-pages, auto updates with each commit
    + Update in settings the location
* Badges; Maintained, Life Cycle, License, Issues, etc
    + https://github.com/GuangchuangYu/badger
        - Commands to create badges

badger package creates the text for pkg readme file.
        
Also
https://shields.io/        
https://github.com/badges/shields
Can make some badges dynamic, static ones below

# issues
[![GitHub issues](https://img.shields.io/github/issues/leppott/MBSStools.svg)](https://GitHub.com/leppott/MBSStools/issues/)
# release
[![GitHub release](https://img.shields.io/github/release/leppott/MBSStools.svg)](https://GitHub.com/leppott/MBSStools/releases/)
# downloads
[![Github all releases](https://img.shields.io/github/downloads/leppott/MBSStools/total.svg)](https://GitHub.com/leppott/MBSStools/releases/)
        
```{r}
pkg_GH <- "leppott/LakeMonitoR"

# GitHub Action, CI
usethis::use_github_actions()
usethis::use_github_actions_badge()
# badger::badge_github_actions(pkg_GH)

# Code coverage
usethis::use_github_action("test-coverage")

# GitHub Action, pkgdown
usethis::use_github_action("pkgdown")

# Badges
# lifecycle
badger::badge_lifecycle("stable", "green") # stable/green, maturing/dormant/blue
# maintained
# license
badger::badge_license() # blank gets from DESCRIPTION
# code quality
badger::badge_codefactor(pkg_GH)
# code testing coverage
badger::badge_codecov(pkg_GH)
# cmd check
badger::badge_github_actions(pkg_GH)

```
